---
title: "Reproducible Data Project 1"
author: "Leonard"
date: "7/3/2020"
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: inline
---
============================================
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  fig.path = "README_figs/README-")
```


## Introduction
This document is submitted for Coursera Reproducible Research Project I.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The data for this assignment can be downloaded from the course web site:

[Activity Monitoring data [52K]](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

The variables included in this dataset are:

* **steps**: Number of steps taking in a 5-minute interval (missing values are coded as **NA**)

* **date**: The date on which the measurement was taken in YYYY-MM-DD format

* **interval**: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

## Preparing The Data
To download the data to your working directory, use `download.file()` and name the destfile according to your desire. After that, unzip the file using the `unzip` function
```{r download file, results=FALSE, cache=TRUE}
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if(!file.exists("activity.csv")){
download.file(url, destfile = "./dataproj1.zip")
unzip("./dataproj1.zip")}
```

## Reading and Processing The Data
After the unzipped file is in your working directory, load it into R and assign it into variable using the `fread` function from `data.table` package. Don't forget to assign "NA" as the `na.strings` argument.

```{r}
library(data.table)
dataRaw <- fread("./activity.csv", na.strings = "NA")
```

Check the data information using simple function such as `head`, `dim`, and `str`

Using dim to check the dimensions of the data.
```{r}
dim(dataRaw)
```
```
Using head to check the upper rows of the data.
```{r}
head(dataRaw)
```

Using str to see the strucutre of the data.
```{r}
str(dataRaw)
```


Since the date column isn't in `date` format, we shall change it into `date` format using `as.Date` function.
```{r}
dataRaw$date <- as.Date(dataRaw$date, format = "%Y-%m-%d")
```

Before analyzing the data, we will remove the NA values and take only the complete data. Since analysis will be split into multiple part (With or Without Removing data), we shall create 2 data to fulfill this demand.

```{r}
WONAdata <- dataRaw[complete.cases(dataRaw), ] ## this will be the data without NA, 

```


## PART I: What is mean total number of steps taken per day?

To make this plot, we'll use base plotting system. First, **calculate steps based on date** using `aggregate` function and assign it into variable named `agg`.
```{r Sum steps per day}
agg1 <- aggregate(steps ~ date, data = WONAdata, FUN = sum)
head(agg1)
```

After that, simply call `hist` function and assign steps as the argument.
```{r}
hist(agg1$steps, main = "Histogram of The Total Number Of Steps Taken Each Day")

```

Next we will be calculating mean and median of the total number of steps taken per day. 

Simply call `mean` and `median on our aggregated data.

**Calculating Mean**
```{r}
mean(agg1$steps)
```

**Calculating Median**
```{r}
median(agg1$steps)
```


## PART II: What is the average daily activity pattern?

To plot a time series of average number of steps taken, we can use `aggregate` function to calculate average steps taken across all days based on `interval` variable. 

```{r time series}
agg3 <- aggregate(steps ~ interval, data = WONAdata, mean)
```

Plot `agg3` using `ggplot` and line as the plotting method.
```{r TS}
library(ggplot2)
ggplot(agg3, aes(x=interval, y= steps)) + geom_line() + labs(title = "Time Series of Average Steps Taken Across All Days", x = "Interval in Minutes", y= "Average of Steps")
```

To answer which 5-minute interva has the maximum average value, simply use the `which.max` function as follows.
```{r}
paste ("The Answer is the ",
as.character(agg3$interval[which.max(agg3$steps)]) 
        ,"-minute interval" , sep ="")
```


## PART III: Imputing Missing Values
Note that there are a number of days/intervals where there are missing values `NA`. The presence of missing days may introduce bias into some calculations or summaries of the data.

First, we shall check the number of missing values in each column using these following lines.
```{r}
summary(dataRaw)
```

To deal with missing values, we simply assign average values of steps taken on that day to fill the missing steps value.

Since we'll be comparing with and without NA's data, recall that `dataRaw` variable is the variable with NA values in it, To avoid confusion, let's create a new variable `imputeData` , this will be our data with NA values filled.
```{r adding missing values}
imputeData <- dataRaw
for (i in 1:nrow(imputeData)) {
    if(is.na(imputeData$steps[i])){
        imputeData$steps[i] <- agg3$steps[agg3$interval == imputeData$interval[i]] 
    }
}
```

Let's view some of the steps column on this data.
```{r}
print(imputeData)
```


Lets make a histogram of steps taken each day using this newly data.
```{r}
par(mfrow = c(1,2))
agg4 <- aggregate(steps ~ date, data = imputeData, sum)

hist(agg4$steps, main = "Histogram of Total Steps \n Taken Each Day \n with NA Values Replaced", ylim=c(0,35),  xlab = "Steps")

hist(agg1$steps, main = "Histogram of Total Steps \n Taken Each Day \n with NA Values Removed",  ylim=c(0,35), xlab = "Steps")
```

For the **mean** and **median**, we'll again use the aggregate function to calculate the values.
```{r}
agg5 <- aggregate(steps ~ date, data = imputeData, FUN = sum)
```

After having the calculated data for, we shall calculate both **mean** and **median**.

Calculating **Mean**
```{r}
mean(agg1$steps)
```

**Calculating Median**
```{r}
median(agg1$steps)
```

As you can see from the results, imputing NA values didn't affect the median and mean value significantly. 

## PART IV: Are there differences in activity patterns between weekdays and weekends?

For this part the `weekdays()` function may be of some help here. Use the dataset with the filled-in missing values for this part.

First, change the date column data into date format.
```{r}
imputeData$date <- as.Date(imputeData$date) 
```

Second determine whether date is weekdays or weekend.
```{r}
library(lubridate)
imputeData$day <- weekdays(imputeData$date)
for(i in 1:nrow(imputeData)){
        if(imputeData[i, ]$day %in% c("Saturday", "Sunday")){
                imputeData[i, ]$day <- "Weekend"
        }
        else {
         imputeData[i, ]$day <- "Weekday"
        }
}
```

Lastly, we shall create another variable to store average steps taken based on two parameters, `date` and `day`.
```{r}
agg6 <- aggregate(steps ~ date + day, imputeData, mean)
```

After that, plot the data using `ggplot` plotting system as follows.
```{r}
ggplot(agg6, aes(x = date, y =steps)) + facet_wrap(.~day) + geom_line(color = "blue") + theme_light()

```

